How to select predictive models for decision-making or causal inference?

### 2024-10-20 how_to_select_latex.tex
iter

Increasingly complex prediction models integrating richer medical data
have flourished because their predictions can be easily
demonstrated and validated on left-out data. But using them to underpin a
decision on whether to treat or not requires more careful validation,
using a metric accounting for the
putative intervention, the $R\text{-risk}$. On average, the $R\text{-risk}$
brings a sizeable benefit to select the most adequate model, even when
model development is based on treated and
untreated population with little differences, as in RCTs. 
%
Our conclusions are that without prior knowledge, the R-risk is a good
default. However, there is much remaining variation, and the R-risk will
not be optimal for every situation. We have identified one such specific
situation: when the causal effect is large compared to the variation of
the baseline effect.

>>>>

Increasingly complex prediction models integrating richer medical data
have flourished because their predictions can be easily
demonstrated and validated on left-out data. But using them to underpin a
decision on whether to treat or not requires more careful validation,
using a metric accounting for the
putative intervention, the $R\text{-risk}$. On average, the $R\text{-risk}$
brings a sizeable benefit to select the most adequate model, even when
model development is based on treated and
untreated population with little differences, as in RCTs. 
%
Our conclusions are that without prior knowledge, the R-risk is a good
default. However, there is much remaining variation, and the R-risk will
not be optimal for every situation. We have identified one such specific
situation: when the causal effect is large compared to the variation of
the baseline effect where the $\tau\text{-risk}_{IPW}$ should be better.
